---
title: "Initial Korean covid 19"
author: "Roselyn Villatoro"
date: "10/10/2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

# Methods
Result - Categorical
Sex - Categorical

Since result and sex are both categorical, we use the cat~cat helper file.

$H_0$:Dying is independent of sex.    
$H_A$:Dying is dependent of sex.    


## Graphical results - use ggplot commands to do this  


```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

  
The bar chart shows that men die at a higher rate compared to women.




  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Among the people who died there was 64% of men who died and 36% of women who died.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value from the chi-square test was 0.001. So we reject the null hypothesis because the p-value is less than 0.05.

## Conclusions

We conclude that dying is dependent of sex, so we reject the null hypothesis because the p-value from the chi-square test was 0.001 which is less than 0.05. Therefore in the fisher's exact test, it shows that men are 2.3 times more likely to die from Covid in Korea than women in Korea. 

